The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-06-09 manuscript.Rmd
[fix #24] include prop of species collected since list publication

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio and OBIS databases for all records they included in the same geographic boundaries as the lists. After filtering these lists to only include the same higher taxonomic level taxa as the species lists, and keeping only records for which the species could be matched to a species name in WoRMS, we had for the Gulf of Mexico `r nrow(idigbio_gom_records)` and `r nrow(obis_gom_records)` records for iDigBio and OBIS respectively, representing `r dplyr::n_distinct(idigbio_gom_records$worms_valid_name)` and `r dplyr::n_distinct(obis_gom_records$worms_valid_name)` species. For the Pacific Northwest, we retrieved `r nrow(idigbio_kozloff_records)` and `r nrow(obis_kozloff_records)` records for iDigBio and OBIS respectively representing `r dplyr::n_distinct(idigbio_kozloff_records$worms_valid_name)` and `r dplyr::n_distinct(obis_kozloff_records$worms_valid_name)` species. For the Gulf of Mexico, the majority of the species included in the expert lists are recorded in both iDigBio and OBIS (`r format_output(species_in_common$n_spp_gom_idigbio_obis/species_in_common$n_spp_gom*100)` %), or only OBIS (`r format_output(species_in_common$n_spp_gom_obis/species_in_common$n_spp_gom*100)` %) (Fig. \@ref(fig:database-overlap)). For the Pacific Northwest, these proportions were much lower with only `r format_output(species_in_common$n_spp_kozloff_idigbio_obis/species_in_common$n_spp_kozloff*100)`% of the species from the regional list had records in OBIS and iDigBio, and `r format_output(species_in_common$n_spp_kozloff_obis/species_in_common$n_spp_kozloff*100)`%  in OBIS only (Fig. 1). Overall, there is more overlap among the data sources for the Gulf of Mexico than the Pacific Northwest.

@@

```{r compare_database_overlap, fig.cap='Comparison of species by phylum across the public databases and the faunistic list.', ech=FALSE}
knitr::include_graphics("figures/database_overlap.pdf")
```

Species that are included in the list but do not have records in databases could be (1) rare: they have never been collected or have been collected only a few times, and the corresponding records have not been digitized yet; (2) difficult to preserve: large, very small, or fragile species are preserved less frequently; (3) are in the database under a different name: the species has been synonimized or re-identified since the publication of the list under a different name that may be included in the database; (4) at the limit of its range: it was included in the list because it is known to occur at the limit of its geographic zone, but has never been collected within the boundaries.

Species that are in the database but not in the list could be: (1) records not considered by the authors of the checklist: either because they were not accessible or known to the authors; (2) records erroneously excluded the records (2) records that have been discovered since the list has been assembled; (3) records with incorrect identification; (4) records with incorrect geographic coordinates: their associated geographic information in the database places them in the Gulf of Mexico or the Pacific Northwest, but they were actually collected elsewhere.

>>>>

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio and OBIS databases for all records they included in the same geographic boundaries as the lists. After filtering these lists to only include the same higher taxonomic level taxa as the species lists, and keeping only records for which the species could be matched to a species name in WoRMS, we had for the Gulf of Mexico `r nrow(idigbio_gom_records)` and `r nrow(obis_gom_records)` records for iDigBio and OBIS respectively, representing `r dplyr::n_distinct(idigbio_gom_records$worms_valid_name)` and `r dplyr::n_distinct(obis_gom_records$worms_valid_name)` species. For the Pacific Northwest, we retrieved `r nrow(idigbio_kozloff_records)` and `r nrow(obis_kozloff_records)` records for iDigBio and OBIS respectively representing `r dplyr::n_distinct(idigbio_kozloff_records$worms_valid_name)` and `r dplyr::n_distinct(obis_kozloff_records$worms_valid_name)` species. For the Gulf of Mexico, the majority of the species included in the expert lists are recorded in both iDigBio and OBIS (`r format_output(species_in_common$n_spp_gom_idigbio_obis/species_in_common$n_spp_gom*100)` %), or only OBIS (`r format_output(species_in_common$n_spp_gom_obis/species_in_common$n_spp_gom*100)` %) (Fig. \@ref(fig:database-overlap)). For the Pacific Northwest, these proportions were much lower with only `r format_output(species_in_common$n_spp_kozloff_idigbio_obis/species_in_common$n_spp_kozloff*100)`% of the species from the regional list with records in OBIS and iDigBio, and `r format_output(species_in_common$n_spp_kozloff_obis/species_in_common$n_spp_kozloff*100)`%  in OBIS only (Fig. \@ref(fig:database-coverage)).

@@

We investigated the number of species for each phylum that contributed to the discrepancies between the regional species lists and the database records. We found that the better studied phyla contributed the most to the differences across the data sources. OBIS tends to be closer to the full record set, but iDigBio has more species for the Gulf of Mexico that are not included in the list (Fig. \@ref(fig:database-overlap))

Overall, there is more overlap among the data sources for the Gulf of Mexico than the Pacific Northwest.

```{r compare_database_overlap, fig.cap='Comparison of species by phylum across the public databases and the faunistic list. \\#fig:database-overlap', ech=FALSE}
knitr::include_graphics("figures/database_overlap.pdf")
```

Species that are included in the list but do not have records in databases could be (1) rare: they have never been collected or have been collected only a few times, and the corresponding records have not been digitized yet; (2) difficult to preserve: large, very small, or fragile species are preserved less frequently; (3) are in the database under a different name: the species has been synonimized or re-identified since the publication of the list under a different name that may be included in the database; (4) at the limit of its range: it was included in the list because it is known to occur at the limit of its geographic zone, but has never been collected within the boundaries; (5) does not have geographic coordinates associated with it: it can be in the database but without georeferenced datapoints that made it possible to discover them.

Species that are in the database but not in the list could be: (1) records not considered by the authors of the checklist: either because they were not accessible or known to the authors; (2) records erroneously excluded the records (2) records that have been discovered since the list has been assembled (for iDigbio records from the Gulf of Mexico, this represent `r format_output(mean(recently_collected$gom_in_idig$min_year > 2001))`%, n=`r nrow(recently_collected$gom_in_idig)` and `r format_output(mean(recently_collected$pnw_in_idig$min_year > 1987))`%, n=`r nrow(recently_collected$pnw_in_idig)`); (3) records with incorrect identification; (4) records with incorrect geographic coordinates: their associated geographic information in the database places them in the Gulf of Mexico or the Pacific Northwest, but they were actually collected elsewhere.
